SLE martingales in coset conformal field theory

### 2012-04-25 AffineLieAlgebras/SLE/paper.tex
Some changes in SLE paper. Added Smirnov-Makarov paper to bibliography.

\abstract{
  In order to provide a way for the concurrent use of Schramm-Loewner evolution (SLE) and conformal field theory (CFT) we consider CFT correlation functions which are martingales with respect to SLE.  We establish necessary and sufficient conditions on such observables for coset models of conformal field theory and Schramm-Loewner evolution on group manifolds and check the consistency with the  behaviour of parafermionic models. Both SLE and CFT are popular instruments to study critical behavior of physical models in two dimensions, but they use different objects. While Scharmm-Loewner evolution has natural connection with lattice models and is suitable for strict proofs, it lacks computational and predictive power of conformal field theory.}

@@

Schramm-Loewner evolution \cite{schramm2000scaling} is a stochastic process used to study critical behavior \cite{Cardy:2005kh,rohde2005basic}. For many lattice models  the convergence of  domain walls to SLE traces was proved \cite{bauer20062d,schramm2006conformally}.

CFT on the other hand is formulated in terms of families of quantum fields which usually has no natural interpretation in microscopic terms. On the other hand CFT has efficient computational tools such as Virasoro symmetry, Knizhnik-Zamolodchikov equations, fusion algebra etc and provides numerous predictions, such as celebrated Cardy formula which gives crossing probability for critical percolation.

Combination of these two approaches should be immensely powerful, so it is important to find objects which can be studied from both points of view. Correlation functions of CFT which are martingales with respect to Schramm-Loewner evolution are such objects. They are actively studied.

We present new class of such observables. We consider correlation functions in coset conformal field theory which are martingales with respect to Schramm-Loewner evolution with additional random motion on factor of group $G$ by subgroup $A$. Using Knizhnik-Zamolodchikov equations we establish necessary and sufficient conditions on such functions and derive relations connecting parameters of SLE and CFT. We then check the consistency of these results with the earlier results obtained for parafermionic observables and general Cardy conditions on boundary states in conformal field theory.

@@

We
consider Schramm-Loewner evolution $SLE_{\kappa}$ with additional Brownian motion on coset space
$G/A$ characterised by parameter $\tau$ proposed in our paper \cite{2011arXiv1112.4354N}. We consider correlation functions in coset models of
conformal field theory which satisfy martingale conditions with respect to this stochastic process.
Since all minimal unitary models of conformal field theory can be obtained from coset construction
for certain choice of Lie group $G$ and subgroup $A$ \cite{difrancesco1997cft} these correlation functions correspond to SLE
observables appearing as the scaling limit of lattice observables. Trace of
Schramm-Loewner $\gamma_{t}$ evolution describes critical domain wall in lattice models.

 Consider upper half-plane $\mathbb{H}$ with the cut along a critical interface $\gamma_{t}$. We denote this slit domain by $\mathbb{H}_{t}=\mathbb{H}\setminus \gamma_{t}$. Conformal map from $\mathbb{H}_{t}$ to $\mathbb{H}$ is denoted by $g_{t}:\mathbb{H}_{t}\to \mathbb{H}$ and satisfies stochastic differential equation \cite{schramm2000scaling}:
\begin{equation}
\label{eq:19}
  \frac{\partial g_t(z)}{\partial t} = \frac{ 2}{g_t(z)-\sqrt{\kappa}\xi_{t}} ,
\end{equation}
where $\xi_{t}$ is the Brownian motion. The dynamic of the tip $z_{t}$ of critical curve $\gamma_{t}$ (tip of SLE trace) is given by the law $z_{t}=g_{t}^{-1}(\sqrt{\kappa}\xi_{t})$.

@@

It is natural to study the problem in upper half-plane  and consider correlation
functions of boundary conformal field theory with boundary condition changing operators sitting at
the tip of SLE trace and at the infinity.

So we are to consider $N$-point boundary correlation functions on the upper half-plane of the form 
\begin{equation}
  \mathcal{F}(\left\{z_{i},\bar z_{i}\right\}_{i=1}^{N})_{\mathbb{H}_{t}}=
\left<\varphi(z_{t}) \phi_{1}(z_{1},\bar z_{1}) \dots \phi_{n}(z_{n},\bar z_{n})
      \varphi(\infty)\right>,
    \label{eq:2}
\end{equation}
where $\phi_{i}$ are bulk primary fields and $\varphi(z_{t})$ is boundary condition changing operator. We assume conformal invariance of boundary conditions so boundary states satisfy Cardy conditions \cite{cardy1989boundary,cardy1986effect,cardy1984conformal}.

 We consider Schramm-Loewner evolution with additional Brownian motion on $G/A$-coset space \cite{2011arXiv1112.4354N}. Denote by $\mathcal{K}$ Killing form of Lie algebra $\gf$ corresponding to Lie group $G$ and by $t^{a}_{i}$ ($\tilde{t}^{b}_{i}$) the  generators of $\gf$-representation ($\af$-representation) corresponding to primary field $\phi_{i}$.
Primary fields in coset models are labeled by pairs $(\lambda,\eta)$ of weights $\gf$ and $\af$-representations. For each $\lambda$ possible values of $\eta$ are those which appear in decomposition of representation $L^{(\mu)}_{\gf}=\bigoplus_{\eta} H^{\lambda}_{\eta}\otimes L^{(\eta)}_{\af}$. Some pairs are equivalent, we discuss and use this equivalence relation later. 
We assume that boundary field $\varphi$ is also primary and labeled by weights $(\mu,\nu)$.

  We use the conformal map  $w(z):\mathbb{H}\setminus\gamma_{t}\to \mathbb{H}$ to rewrite expression \eqref{eq:2} in the whole upper half plane:

\begin{multline}
  \mathcal{F}(\left\{z_{i},\bar z_{i}\right\})_{\mathbb{H}_{t}}=\\
  =\prod \left(\frac{\partial w(z_{i})}{\partial z_{i}}\right)^{h_{i}} 
  \prod \left(\frac{\partial \bar w(\bar z_{i})}{\partial \bar z_{i}}\right)^{\bar h_i}
  \mathcal{F}(\left\{w_{i}, \bar w_{i}\right\})_{\mathbb{H}}
  \label{eq:1}
\end{multline}
We can extend the definition of $\mathcal{F}$ to the whole plane by doubling the number of fields $\phi_{i}$ \cite{cardy2004boundary,cardy1984conformal} and consider $w_{i},\bar w_{i}$ as independent variables. To simplify notation we will write $\mathcal{F}(\left\{w_{i}\right\}_{i=1}^{2N})$.

@@

We denote by $\mathcal{G}_{i}$ generator of infinitesimal transformation of primary field $\phi_{i}$:$d\phi_{i}(w_{i}) = \mathcal{G}_{i}\phi_{i}(w_{i})$. We normalize additional $\dim\gf$-dimensional Brownian motion as $\mathbb{E}\left[d\theta^{a}\; d\theta^{b}\right]=\mathcal{K}(t^{a},t^{b})dt$, and the generator of field transformation \eqref{eq:2} is equal to 

@@

 Ito formula  gives us the expression for the differential of $\mathcal{F}$, it should be zero due to martingale condition:

@@

\begin{array}{rl}
  L_{n}=& L_{n}^{\gf}-L_{n}^{\af}\\
  \left[L_{n}^{\gf},J^{a}_{m}\right]=& -m J^{a}_{n+m}\\
  \left[L_{n}^{\gf},\tilde{J}^{b}_{m}\right]=&-m\tilde{J}^{b}_{n+m}\\
  \left[L_{n}^{\gf},L_{m}\right]=&[L_{n}^{\gf},L_{m}^{\gf}]-[L_{n}^{\gf},L_{m}^{\af}]=\\
  & =(n-m)L_{m+n}+\frac{c}{12}(n^{3}-n)\delta_{m+n,0}

@@

@@

\section{Outlook}
\label{sec:outlook}

In excellent paper \cite{bettelheim2005stochastic} connection between WZNW-models and Schramm-Loewner evolution with additional Brownian motion on group manifold was established. This paper also stated a problem of the connection of SLE parameters for martingales in WZNW, coset and minimal models. Here we used the method to obtain necessary and sufficient conditions on SLE martingales proposed in paper \cite{alekseev2010sle}. This method allowed us to compare our results with parafermionic results of the papers \cite{santachiara2008sle,picco2008numerical}.

Coset structure of minimal models is manifest in field theory perturbed by external magnetic field \cite{fateev1990conformal,eguchi1989deformations,hollowood1989rational}. This theory is supported by experimental data \cite{coldea2010quantum}. Relation between correlation functions in coset theory and SLE observables can be a starting point for the study of domain walls in lattice models away of critical point. Schramm-Loewner evolution was applied for the description of domain walls in Ising model with external field in recent paper \cite{stevenson2011domain}, although the perturbation was not compatible with affine field theory structure.

\section{Conclusion}
\label{sec:conclusion}

>>>>

\abstract{
  Scharmm-Loewner evolution (SLE) and conformal field theory (CFT) are popular instruments to study critical behavior of physical models in two dimensions, but they use different objects. While SLE  has natural connection with lattice models and is suitable for strict proofs, it lacks computational and predictive power of conformal field theory.
  To provide a way for the concurrent use of SLE and CFT we consider CFT correlation functions which are martingales with respect to SLE.  We establish necessary and sufficient conditions on such observables for coset models of conformal field theory and Schramm-Loewner evolution on group manifolds and check the consistency with the  behaviour of parafermionic models. Coset models are connected with off-critical massive field theories and we discuss possible implications for SLE.}

@@

Schramm-Loewner evolution \cite{schramm2000scaling} is a stochastic process satisfying stochastic differential equation (\ref{eq:19}). Solution of equation (\ref{eq:19}) can be seen as probability measure on random curves called SLE traces. SLE is  useful for the study of critical behavior \cite{Cardy:2005kh,rohde2005basic}, since for many lattice models  the convergence of  domain walls to SLE traces can be proved \cite{bauer20062d,schramm2006conformally}. Practical computations with SLE rely on Ito calculus.

Conformal field theory \cite{belavin1984ics} is formulated in terms of families of quantum fields which usually has no natural interpretation in microscopic terms. On the other hand CFT has efficient computational tools such as Virasoro symmetry, Knizhnik-Zamolodchikov equations, fusion algebra etc and provides numerous predictions, such as celebrated Cardy formula \cite{cardy1992critical,smirnov2001critical} which gives crossing probability for critical percolation.

Combination of these two approaches should be immensely powerful, so it is important to find objects which can be studied from both points of view. We consider correlation functions of CFT which are martingales with respect to Schramm-Loewner evolution are such objects.

We present new class of such observables in coset conformal field theory \cite{Goddard198588}. They are correlation functions in $G/A$-coset conformal field theory which are martingales with respect to Schramm-Loewner evolution with additional Brownian motion on factor-space of Lie group $G$ by subgroup $A$ \cite{2011arXiv1112.4354N}. Using Knizhnik-Zamolodchikov equations we establish necessary and sufficient conditions on such functions and derive relations connecting parameters of SLE and CFT. We then check the consistency of these results with the earlier results obtained for parafermionic observables \cite{santachiara2008sle}.

All minimal unitary models can be obtained by coset construction. This coset structure leads to specific off-critical behavior such as classification of massive excitations by affine Toda field theory \cite{fateev1990conformal,eguchi1989deformations,hollowood1989rational}. In conclusion we discuss possible connections of these CFT perturbation with massive off-critical SLEs (which contain additional drift term  \cite{makarov2010off}).

@@

%% We
%% consider Schramm-Loewner evolution $SLE_{\kappa}$ with additional Brownian motion on coset space
%% $G/A$ characterised by parameter $\tau$ proposed in our paper \cite{2011arXiv1112.4354N}. We consider correlation functions in coset models of
%% conformal field theory which satisfy martingale conditions with respect to this stochastic process.
%% Since all minimal unitary models of conformal field theory can be obtained from coset construction
%% for certain choice of Lie group $G$ and subgroup $A$ \cite{difrancesco1997cft} these correlation functions correspond to SLE
%% observables appearing as the scaling limit of lattice observables. Trace of
%% Schramm-Loewner $\gamma_{t}$ evolution describes critical domain wall in lattice models.
%% 
 Consider some critical lattice model on the upper half-plane $\mathbb{H}$ with the cut along a critical interface $\gamma_{t}$. We denote this slit domain by $\mathbb{H}_{t}=\mathbb{H}\setminus \gamma_{t}$.  Assume that $\gamma_{t}$ satisfies restriction property and is conformally  invariant \cite{Cardy:2005kh}. Then  conformal map  $g_{t}:\mathbb{H}_{t}\to \mathbb{H}$ satisfies stochastic differential equation \cite{schramm2000scaling}:
\begin{equation}
\label{eq:19}
  \frac{\partial g_t(z)}{\partial t} = \frac{ 2}{g_t(z)-\sqrt{\kappa}\xi_{t}} ,
\end{equation}
where $\xi_{t}$ is the Brownian motion. The dynamic of the tip $z_{t}$ of critical curve $\gamma_{t}$ (tip of SLE trace) is given by the law $z_{t}=g_{t}^{-1}(\sqrt{\kappa}\xi_{t})$.

@@

Consider  $N$-point correlation function of boundary conformal field theory with boundary condition changing operators sitting at the tip of SLE trace and at the infinity:
\begin{equation}
  \mathcal{F}(\left\{z_{i},\bar z_{i}\right\}_{i=1}^{N})_{\mathbb{H}_{t}}=
\left<\varphi(z_{t}) \phi_{1}(z_{1},\bar z_{1}) \dots \phi_{n}(z_{n},\bar z_{n})
      \varphi(\infty)\right>,
    \label{eq:2}
\end{equation}
where $\phi_{i}$ are bulk primary fields and $\varphi(z_{t})$ is boundary condition changing operator.

Primary fields in coset models are labeled by pairs $(\lambda,\eta)$ of weights $\gf$ and $\af$-representations. For each $\lambda$ possible values of $\eta$ are those which appear in decomposition of representation $L^{(\mu)}_{\gf}=\bigoplus_{\eta} H^{\lambda}_{\eta}\otimes L^{(\eta)}_{\af}$. Some pairs are equivalent, we discuss and use this equivalence relation later. 
We assume that boundary field $\varphi$ is also primary and labeled by weights $(\mu,\nu)$.

  We use the conformal map  $w(z):\mathbb{H}\setminus\gamma_{t}\to \mathbb{H}$ to rewrite expression \eqref{eq:2} in the whole upper half plane:
\begin{multline}
  \mathcal{F}(\left\{z_{i},\bar z_{i}\right\})_{\mathbb{H}_{t}}=\\
  =\prod \left(\frac{\partial w(z_{i})}{\partial z_{i}}\right)^{h_{i}} 
  \prod \left(\frac{\partial \bar w(\bar z_{i})}{\partial \bar z_{i}}\right)^{\bar h_i}
  \mathcal{F}(\left\{w_{i}, \bar w_{i}\right\})_{\mathbb{H}}
  \label{eq:1}
\end{multline}
We can extend the definition of $\mathcal{F}$ to the whole plane by doubling the number of fields $\phi_{i}$ \cite{cardy2004boundary,cardy1984conformal} and consider $w_{i},\bar w_{i}$ as independent variables. To simplify notation we will write $\mathcal{F}(\left\{w_{i}\right\}_{i=1}^{2N})$.

$G/A$-coset conformal field theory can be realized as WZNW-model with gauge group $G$ interacting with pure gauge field of gauge group $A\subset G$ \cite{gawdzki1988g,figueroa89equivalence}. So we have $G/A$ gauge symmetry and we add random gauge transformations to Schramm-Loewner evolution \cite{2011arXiv1112.4354N} similar to the case of WZNW-models \cite{bettelheim2005stochastic}. Denote by $\mathcal{K}$ Killing form of Lie algebra $\gf$ corresponding to Lie group $G$ and by $t^{a}_{i}$ ($\tilde{t}^{b}_{i}$) the  generators of $\gf$-representation ($\af$-representation) corresponding to primary field $\phi_{i}$.

@@

We denote by $\mathcal{G}_{i}$ generator of infinitesimal transformation of primary field $\phi_{i}$:$d\phi_{i}(w_{i}) = \mathcal{G}_{i}\phi_{i}(w_{i})$. We normalize additional $\dim\gf$-dimensional Brownian motion as $\mathbb{E}\left[d\theta^{a}\; d\theta^{b}\right]=\mathcal{K}(t^{a},t^{b})dt$, and the generator of field transformation is equal to 

@@

 Ito formula  gives us the expression for the differential of $\mathcal{F}$ which should be zero due to martingale condition:

@@

\begin{array}{ll}
  L_{n}= L_{n}^{\gf}-L_{n}^{\af} &\\
  \left[L_{n}^{\gf},J^{a}_{m}\right]= -m J^{a}_{n+m}&\\
  \left[L_{n}^{\gf},\tilde{J}^{b}_{m}\right]=-m\tilde{J}^{b}_{n+m}&\\
  \left[L_{n}^{\gf},L_{m}\right]=[L_{n}^{\gf},L_{m}^{\gf}]-[L_{n}^{\gf},L_{m}^{\af}]=&\\
  \quad =(n-m)L_{m+n}+\frac{c}{12}(n^{3}-n)\delta_{m+n,0}&

@@

Consider standard coset realization of minimal models: $su(2)_{N}/u(1)_{N}$. Central charge is $c=\frac{3N}{N+2}-1=\frac{2N-2}{N+2}$. Conformal weights of primary fields with Dynkin indeces $(k,l)$ are $h_{(k,l)}=\frac{k(k+2)}{4(N+2)}-\frac{l^{2}}{4N}$. $C_{\mu}=k(k+2)/2, C_{\nu}=l^{2}/2$
For Ising model $N=2,c=1/2$, $h_{(2,0)}=1/2, h_{(1,1)}=1/16$. We get for $\varphi_{(2,0)}$: $3\kappa-9+4\tau =0, -3+\kappa+4\tau=0$, $\kappa=3, \tau=0$. For $\varphi_{(1,1)}$: $3\kappa-16+64\tau=0, -64+9\kappa + 64\tau=0$, $\kappa=16/3, \tau=0$. 
For $N=3, c=4/5$. Conformal weights $h_{(0,0)}=0$, $h_{(0,2)}=2/3$, $h_{(0,-2)}=2/3$, $h_{(1,3)}=2/5$, $h_{(1,1)}=1/15$, $h_{(1,-1)}=1/15$.  Corresponding values of $\kappa,\tau$ are: $(208/25,242/225), (460/33,116/33),(260/53,28/477)$

Now use another coset realization of minimal unitary models: $\frac{\hat{su}(2)_{N}\oplus \hat{su}(2)_{1}}{\hat{su}(2)_{N+1}}$, $c=\frac{3N}{N+2}+1-\frac{3(N+1)}{N+3}=1-\frac{6}{(N+2)(N+3)}$

three-state Potts model $\frac{\hat{su}(3)_{1}\oplus \hat{su}(3)_{1}}{\hat{su}(3)_{2}}$

@@

\section{Conclusion and outlook}
\label{sec:outlook}

In excellent paper \cite{bettelheim2005stochastic} connection between WZNW-models and Schramm-Loewner evolution with additional Brownian motion on group manifold was established. This paper also stated a problem of the connection of SLE parameters for martingales in WZNW, coset and minimal models. In present letter we used the method to obtain necessary and sufficient conditions on SLE martingales proposed in paper \cite{alekseev2010sle}. This method allowed us to compare our results with parafermionic results of the papers \cite{santachiara2008sle,picco2008numerical}.

Coset structure of minimal models is manifest in field theory perturbed by external magnetic field \cite{fateev1990conformal,eguchi1989deformations,hollowood1989rational}. This theory is supported by experimental data \cite{coldea2010quantum}. Relation between correlation functions in coset theory and SLE observables can be a starting point for the study of domain walls in lattice models away of critical point.

Massive off-critical SLEs has additional drift term in driving Brownian motion \cite{makarov2010off}.

  Schramm-Loewner evolution was applied for the description of domain walls in Ising model with external field in recent paper \cite{stevenson2011domain}, although the perturbation was not compatible with affine field theory structure.
